---
title: 'Data Exploration: Candy Study Dataset'
output:
  html_document:
    df_print: paged
---


#Load libraries 
```{r}
# Libraries need to be installed before they can be used (but only once)

#install.packages("magrittr")
#install.packages("plyr")
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("grid")
#install.packages("gridExtra")
#install.packages("stringr")
#install.packages("reshape2")


#STRG+SHIFT+RETURN
# load already installed libraries 
library(magrittr)
library(plyr)
library(dplyr)
library(ggplot2)
library(grid)
library(gridExtra)
library(stringr)
library(here)
library(scales)
library(readr)
library(tidyverse)
library(tibble)
library(reshape2)

```


#Data Beschreibung


[Source](https://github.com/fivethirtyeight/data/blob/master/candy-power-ranking/candy-data.csv)
Header	Description
  -chocolate:	Does it contain chocolate?
  -fruity	:   Is it fruit flavored?
  -caramel: 	Is there caramel in the candy?
  -peanutalmondy:	Does it contain peanuts, peanut butter or almonds?
  -nougat:	Does it contain nougat?
  -crispedricewafer:	Does it contain crisped rice, wafers, or a cookie component?
  -hard:	Is it a hard candy?
  -bar:	Is it a candy bar?
  -pluribus:	Is it one of many candies in a bag or box?
  -sugarpercent:	The percentile of sugar it falls under within the data set.
  -pricepercent:	The unit price percentile compared to the rest of the set.
  -winpercent:	The overall win percentage according to 269,000 matchups.




##upload Data
```{r}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"

# Read the CSV data into a data frame
candy_raw <- read.csv(url)
```

# Descriptive Statistics



#Show first 5 rows. 
```{r}
head(candy_raw)
```

#Show structure information about the dataframe. 
```{r}
str(candy_raw)
```

#Show summary statistics per column.
```{r}
summary(candy_raw)
```


# Cleaning and Preparing the Data

Based on the results of the str(candy_raw) function above, several issues with how the read.csv function imported the data were identified that need to be cleaned up before going in-depth with the analysis:



```{r}

## competitorname column is useless for analysis 
candy <- candy_raw[ , -1 ]

## lets spell out the chocolate flag 
candy$chocolate <- ifelse (candy$chocolate == 0, "no chocolate", "chocolate")


## lets spell out the fruity flag 
candy$fruity <- ifelse (candy$fruity == 0, "not fruity", "fruity")

## lets spell out the caramel flag 
candy$caramel <- ifelse (candy$caramel == 0, "no caramel", "caramel")

##lets spell out peanutyalmondy
candy$peanutyalmondy <- ifelse (candy$peanutyalmondy == 0, "not peanutyalmondy", "peanutyalmondy")


##lets spell out the nougat flag 
candy$nougat <- ifelse (candy$nougat == 0, "no nougat", "nougat")

##lets spell out the crispedricewafer flag 
candy$crispedricewafer <- ifelse (candy$crispedricewafer == 0, "no crispedricewafer", "crispedricewafer")

##lets spell out the hard flag 
candy$hard <- ifelse (candy$hard == 0, "not hard", "hard")

##lets spell out the bar flag 
candy$bar <- ifelse (candy$bar == 0, "not bar", "bar")

##lets spell out the pluribus flag 
candy$pluribus <- ifelse (candy$pluribus == 0, "no pluribus", "pluribus")


## rename column gewinpercent
#colnames(candy)[12] <- "target" 




```


# Univariate Plots
In this section I will take a look at the distribution of values for each variable in the dataset by creating histograms using ggplot2's qplot function. We are trying to find out if there is more data to clean up, including outliers or extraneous values. This also might help me begin to identify any relationships between variables that are worth investigating further.



##barchart for top ten 
```{r}


# Sort the candies' dataframe in descending order based on the 'Overall Score' column
sorted_candies <- candy_raw[order(candy_raw$winpercent, decreasing = TRUE),]

# Fetch the top 10 candies from the sorted dataframe
top_10_candies <- head(sorted_candies, 10)

# Print the top 10 candies
print(top_10_candies)

# Reorder levels in ascending order
top_10_candies$competitorname <- factor(top_10_candies$competitorname, levels = unique(top_10_candies$competitorname))

# Create a bar plot
bar_plot <- ggplot(top_10_candies, aes(x = winpercent, y = reorder(competitorname, winpercent))) +
  geom_bar(stat = "identity", color = "black",fill = "orange") +
  labs(title = "Top 10 Competitors Based on Overall Win Percent",
       x = "Win Percent", y = "Competitor Name") +
  theme_minimal() +
  theme(legend.position = "none")  # Remove legend for better appearance

# Display the plot
print(bar_plot)
```




## chococlate Distribution  

```{r}

chocolate_car <- candy %>%
                  ggplot(aes(x = chocolate, fill = chocolate)) + 
                    geom_histogram(stat="count") +
                    theme(text = element_text(size = 16)) +
                    labs(title = "Frequency Histogram: Presence of Chocolate in Candy",
                         x = "Chocolate",
                         y = "Count")
 print(chocolate_car)

# Berechnung der Pearson-Korrelation
chocolate_corr <- cor(candy_raw$chocolate, candy_raw$winpercent)
cat("Pearson-Korrelation:", round(chocolate_corr, 3), "\n")



```
**Interpretation** 
The Graph Tell us that from the Candy sample, the majority never had Chocolate

##fruity Distribution

```{r}

fruity_car <- candy %>%
                ggplot(aes(x = fruity, fill = fruity)) + 
                  geom_histogram(stat="count") +
                  theme(text = element_text(size = 16)) +
                  labs(title = "Frequency Histogram: Where the Candys fruity?",
                       x = "Fruity",
                       y = "Count")

print(fruity_car)


# Berechnung der Pearson-Korrelation
fruity_corr <- cor(candy_raw$fruity, candy_raw$winpercent)
cat("Pearson-Korrelation:", round(fruity_corr, 3), "\n")



```
**Interpretation** 
The Graph Tell us that from the Candy sample, the majority never not Fruity



##caramel Distribution

```{r}

caramel_car <- candy %>%
                ggplot(aes(x = caramel, fill = caramel)) + 
                  geom_histogram(stat="count") +
                  theme(text = element_text(size = 16)) +
                  labs(title = "Frequency Histogram: Caramel?",
                       x = "Caramel",
                       y = "Count")
 print(caramel_car)

# Berechnung der Pearson-Korrelation
caramel_corr <- cor(candy_raw$caramel, candy_raw$winpercent)
cat("Pearson-Korrelation:", round(caramel_corr, 3), "\n")



```
**Interpretation** 
The Graph Tell us that from the Candy sample, the majority never had no Caramel


## Peanutyalmondy Distribution  

```{r}

Peanutyalmondy_car <- candy %>%
                        ggplot(aes(x = peanutyalmondy, fill = peanutyalmondy)) + 
                          geom_histogram(stat="count") +
                          theme(text = element_text(size = 16)) +
                          labs(title = "Frequency Histogram:Presence of Peanuts and Almonds",
                               x = "Chocolate",
                               y = "Count")

print(Peanutyalmondy_car)


# Berechnung der Pearson-Korrelation
peanutyalmondy_corr <- cor(candy_raw$peanutyalmondy, candy_raw$winpercent)
cat("Pearson-Korrelation:", round(peanutyalmondy_corr, 3), "\n")



```
**Interpretation** 
The Graph Tell us that from the Candy sample, the majority never had no Peanuts nor Nuts

##nougat Distribution

```{r}

nougat_car <- candy %>%
              ggplot(aes(x = nougat, fill = nougat)) + 
                geom_histogram(stat="count") +
                theme(text = element_text(size = 16)) +
                labs(title = "Frequency Histogram: Nougat",
                     x = "Nougat",
                     y = "Count")

print(nougat_car)


# Count the number of candies with and without nougat
nougat_counts <- table(candy$nougat)

# Calculate the percentages
percentage_with_nougat <- (nougat_counts["nougat"] / sum(nougat_counts)) * 100
percentage_without_nougat <- (nougat_counts["no nougat"] / sum(nougat_counts)) * 100

# Display the results
cat("Percentage of candies with nougat: ", round(percentage_with_nougat, 2), "%\n")
cat("Percentage of candies without nougat: ", round(percentage_without_nougat, 2), "%\n")


# Berechnung der Pearson-Korrelation
nougat_corr <- cor(candy_raw$nougat, candy_raw$winpercent)
cat("Pearson-Korrelation:", round(nougat_corr, 3), "\n")



```
**Interpretation** 
The Graph Tell us that from the Candy sample, the majority never had no Nougats



##Crispedricewafer Distribution

```{r}

Crispedricewafer_car <- candy %>%
                        ggplot(aes(x = crispedricewafer, fill = crispedricewafer)) + 
                          geom_histogram(stat="count") +
                          theme(text = element_text(size = 16)) +
                          labs(title = "Frequency Histogram: Crispedricewafer",
                               x = "Crispedricewafer",
                               y = "Count")

print(Crispedricewafer_car)


# Scatterplot erstellen
scatter_plot <- ggplot(candy_raw, aes(x = crispedricewafer, y = winpercent, color = crispedricewafer)) +
  geom_point() +  # Punktdiagramm erstellen
  geom_smooth(method = "lm", se = FALSE, color = "blue") + #Regressionline
  scale_color_gradient(low = "blue", high = "red") +  # Farbskala festlegen
  labs(title = "Scatterplot: Sugarpercent vs. Winpercent",
       x = "Sugarpercent",
       y = "Winpercent")

# Plot anzeigen
print(scatter_plot)



# Zählen der Anzahl von Süßigkeiten mit und ohne crispedricewafer
crispedricewafer_counts <- table(candy$crispedricewafer)

# Berechnung der Prozentsätze
prozentsatz_mit_crispedricewafer <- (crispedricewafer_counts["crispedricewafer"] / sum(crispedricewafer_counts)) * 100
prozentsatz_ohne_crispedricewafer <- (crispedricewafer_counts["no crispedricewafer"] / sum(crispedricewafer_counts)) * 100

# Anzeige der Ergebnisse
cat("Prozentsatz der Süßigkeiten mit crispedricewafer: ", round(prozentsatz_mit_crispedricewafer, 2), "%\n")
cat("Prozentsatz der Süßigkeiten ohne crispedricewafer: ", round(prozentsatz_ohne_crispedricewafer, 2), "%\n")


# Berechnung der Pearson-Korrelation
crispedricewafer_corr <- cor(candy_raw$crispedricewafer, candy_raw$winpercent)
cat("Pearson-Korrelation:", round(crispedricewafer_corr, 3), "\n")


```
**Interpretation** 
The Graph Tell us that from the Candy sample, the majority never were not Crispy


##hard Distribution

```{r}

hard_car <- candy %>%
            ggplot(aes(x = hard, fill = hard)) + 
              geom_histogram(stat="count") +
              theme(text = element_text(size = 16)) +
              labs(title = "Frequency Histogram: Hard",
                   x = "Hard",
                   y = "Count")

print(hard_car)


# Erstelle den Scatterplot
ggplot(candy_raw, aes(x = hard, y = winpercent, colour = hard)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") + #Regressionline
  labs(title = "Scatter Plot of Hardness vs. Win Percentage",
       x = "Hardness",
       y = "Win Percentage")

# Zählen der Anzahl von Süßigkeiten mit und ohne hard
hard_counts <- table(candy$hard)

# Berechnung der Prozentsätze
prozentsatz_mit_hard <- (hard_counts["hard"] / sum(hard_counts)) * 100
prozentsatz_ohne_hard <- (hard_counts["not hard"] / sum(hard_counts)) * 100

# Anzeige der Ergebnisse
cat("Prozentsatz der Süßigkeiten mit hard: ", round(prozentsatz_mit_hard, 2), "%\n")
cat("Prozentsatz der Süßigkeiten ohne hard: ", round(prozentsatz_ohne_hard, 2), "%\n")


# Berechnung der Pearson-Korrelation
hard_corr <- cor(candy_raw$hard, candy_raw$winpercent)
cat("Pearson-Korrelation:", round(hard_corr, 3), "\n")


```
**Interpretation** 
The Graph Tell us that from the Candy sample, the majority were not hard


##Bar Distribution

```{r}

bar_car <- candy %>%
            ggplot(aes(x = bar, fill = bar)) + 
              geom_histogram(stat="count") +
              theme(text = element_text(size = 16)) +
              labs(title = "Frequency Histogram: bar",
                   x = "Bar",
                   y = "Count")

print(bar_car)


# Erstelle den Scatterplot
ggplot(candy_raw, aes(x = bar, y = winpercent, colour = bar)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE, color = "blue") + #Regressionline
  labs(title = "Scatter Plot of Hardness vs. Win Percentage",
       x = "Hardness",
       y = "Win Percentage")


# Berechnung der Pearson-Korrelation
bar_corr <- cor(candy_raw$bar, candy_raw$winpercent)
cat("Pearson-Korrelation:", round(bar_corr, 3), "\n")

```
**Interpretation** 
The Graph Tell us that from the Candy sample, the majority never not Bar


##Pluribus Distribution

```{r}

pluribus_car <- candy %>%
                ggplot(aes(x = pluribus, fill = pluribus)) + 
                  geom_histogram(stat="count") +
                  theme(text = element_text(size = 16)) +
                  labs(title = "Frequency Histogram: pluribus",
                       x = "Pluribus",
                       y = "Count")

print(pluribus_car)


# Scatterplot erstellen
scatter_plot <- ggplot(candy_raw, aes(x = pluribus, y = winpercent, color = pluribus)) +
  geom_point() +  # Punktdiagramm erstellen
  geom_smooth(method = "lm", se = FALSE, color = "blue") + #Regressionline
  scale_color_gradient(low = "blue", high = "red") +  # Farbskala festlegen
  labs(title = "Scatterplot: Sugarpercent vs. Winpercent",
       x = "Sugarpercent",
       y = "Winpercent")

# Plot anzeigen
print(scatter_plot)

# Berechnung der Pearson-Korrelation
pluribus_corr <- cor(candy_raw$pluribus, candy_raw$winpercent)
cat("Pearson-Korrelation:", round(pluribus_corr, 3), "\n")




```
**Interpretation** 
The Graph Tell us that from the Candy sample, the ratio of Multi-pieces Candies to non-Multi-pieces Candies were almost the Same

##sugarpercent Distribution

```{r}
sugar_car <- candy %>%
            ggplot(aes(x = sugarpercent, fill = sugarpercent)) + 
              geom_histogram(stat="count") +
              theme(text = element_text(size = 16)) +
              labs(title = "Frequency Histogram: Sugarpercent",
                   x = "Sugarpercent in %",
                   y = "Count")
          

print(sugar_car)


# Scatterplot erstellen
scatter_plot <- ggplot(candy, aes(x = sugarpercent, y = winpercent, color = sugarpercent)) +
  geom_point() +  # Punktdiagramm erstellen
  geom_smooth(method = "lm") + #Regressionline
  scale_color_gradient(low = "blue", high = "red") +  # Farbskala festlegen
  labs(title = "Scatterplot: Sugarpercent vs. Winpercent",
       x = "Sugarpercent",
       y = "Winpercent")

# Plot anzeigen
print(scatter_plot)


# first we need to set up our plot
ggplot(data = candy_raw, aes(x = sugarpercent, 
                                 y = pricepercent, 
                                 label = competitorname)) +
    geom_point() + # then add a layer of points
    geom_smooth(method = "lm") + # and then add a fitted line
    geom_text(check_overlap = T,# automatically reduce overlap (deletes some labels)
              vjust = "bottom", # adjust the vertical orientation
              nudge_y = 0.01, # move the text up a bit so it doesn't touch the points
              angle = 30, # tilt the text 30 degrees
              size = 2 # make the text smaller (to reduce overlap more)
             )



# Berechnung der Pearson-Korrelation
sugar_corr <- cor(candy$sugarpercent, candy$winpercent)
cat("Pearson-Korrelation:", round(sugar_corr, 3), "\n")


```
**Interpretation** 
The Graph Tell us that from the Candy sample, was a positve Correlation between the Sugarpercent and Winpercent. That is an increase in the Sugar percentage in the Candies might likely lead to an increase in the Satisfaction of Customers.



##Pricepercent
```{r}

price_car <- candy %>%
              ggplot(aes(x = pricepercent, fill = pricepercent)) + 
                geom_histogram(stat="count") +
                theme(text = element_text(size = 16)) +
                labs(title = "Frequency Histogram: Pricepercent",
                     x = "Pricepercent in %",
                     y = "Count")

print(price_car)



# Scatterplot erstellen
scatter_plot <- ggplot(candy, aes(x = pricepercent, y = winpercent, color = pricepercent)) +
  geom_point() +  # Punktdiagramm erstellen
  geom_smooth(method = "lm", se = FALSE, color = "blue") + #Regressionline
  scale_color_gradient(low = "blue", high = "red") +  # Farbskala festlegen
  labs(title = "Scatterplot: Pricepercent vs. Winpercent",
       x = "Pricepercent",
       y = "Winpercent")

# Plot anzeigen
print(scatter_plot)


pric_corr <- cor(candy$pricepercent, candy$winpercent)
cat("Korrelation zwischen Sugarpercent und Winpercent:",pric_corr, "\n")





```
**Interpretation** 
The Graph Tell us that from the Candy sample, was a positve Correlation between the Pricepercent and Winpercent. That is an increase in the Price ratio in the Candies might likely lead to an increase in the Satisfaction of Customers.





##Plotting the Correlation Heapmap

```{r}



# Calculate the correlation matrix for numeric columns
correlation_matrix <- cor(candy_raw %>% select_if(is.numeric))

# Create the correlation heatmap
heatmap_plot <- ggplot(data = melt(correlation_matrix), aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  labs(title = "Correlation Heatmap")

# Display the heatmap
print(heatmap_plot)


```




```{r}
# Group the dataset by competitorname and calculate the count
count_data <- top_10_candies %>%
  group_by(competitorname) %>%
  summarise(count = n())

# Sort the data by count in descending order
count_data <- count_data[order(-count_data$count), ]

# Plotting the pie chart
pie(count_data$count, labels = count_data$competitorname, main = "Composition of Highest-Ranked Competitor Name")

# Adding a legend
legend("topright", legend = count_data$competitorname, cex = 0.8, fill = rainbow(length(count_data$competitorname)))

# Adding a title
title(main = "Composition of Highest-Ranked Competitor Name")

# Adjusting the aspect ratio of the plot
options(repr.plot.width = 8, repr.plot.height = 6)

```


##Ploting the Correlation Graph
```{r}


# Create a dataframe from the correlation matrix

 x <- c(bar_corr, caramel_corr, chocolate_corr, crispedricewafer_corr, fruity_corr, hard_corr, nougat_corr, pluribus_corr, pric_corr, sugar_corr, peanutyalmondy_corr)
 
 variables <- c("bar", "caramel", "chocolate", "crispedricewafer", "fruity", "hard", "nougat", "pluribus", "price", "sugar", "peanutyalmondy")
 
 
# Create a tibble
correlation_tibble <- tibble(Variable = variables, Correlation = x)


# Create a colorful bar chart
bar_chart <- ggplot(correlation_tibble, aes(x = reorder(Variable, Correlation), y = Correlation, fill = Variable)) +
  geom_bar(stat = "identity") +
  labs(title = "Correlation between other Variables and the Winpercent",
       x = "Variable", y = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  scale_fill_brewer(palette = "Set3")  # You can choose a different palette

# Display the chart
print(bar_chart)

```

```{r}
# Fit a linear regression model
regression_model <- lm(winpercent ~ sugarpercent + chocolate, data = candy_raw)

# Print the summary of the regression model
summary(regression_model)

```
**interpretation**
This is the output of a linear regression model:

# Residuals:
- The residuals (differences between the observed and predicted values) have a minimum of -23.6981, a maximum of 27.4820, and are centered around 0.

# Coefficients:
- **Intercept (38.262):** The estimated average `winpercent` when `sugarpercent` and `chocolate` are zero.
- **sugarpercent (8.567):** For each unit increase in `sugarpercent`, the `winpercent` is estimated to increase by 8.567 (not statistically significant at conventional levels, p-value = 0.0525).
- **chocolate (18.273):** For each unit increase in `chocolate` (assuming it's a binary variable), the `winpercent` is estimated to increase by 18.273.

# Significance:
- The intercept is highly significant (p-value < 2e-16), indicating that the model is significantly different from a model with no predictors.
- `chocolate` is highly significant (p-value = 1.06e-10), suggesting that it contributes significantly to predicting `winpercent`.

# Goodness-of-Fit:
- **Residual Standard Error (11.22):** The standard deviation of the residuals.
- **Multiple R-squared (0.432):** The proportion of the variance in `winpercent` explained by the model.
- **Adjusted R-squared (0.4181):** The R-squared adjusted for the number of predictors.

# F-statistic:
- **F-statistic (31.18):** A test of the overall significance of the model.
- **p-value (< 2.2e-16):** The model is statistically significant, suggesting that at least one of the predictors is related to `winpercent`.

Overall, the model suggests that both `sugarpercent` and `chocolate` are associated with `winpercent`, and the model provides a good fit to the data.

##Creating a linear regression using Sugarpercent and Fruity to predict winpercent
```{r}
# Create a linear regression model
model <- lm(winpercent ~ sugarpercent + fruity, data = candy_raw)

# Display the summary of the regression model
summary(model)

```
**interpretation**

The output you provided is the summary of a linear regression model. Here's how you can interpret each part:

1. **Call:**
   - It shows the formula used to fit the model. In this case, the model is `winpercent ~ sugarpercent + fruity`, indicating that you're predicting `winpercent` based on `sugarpercent` and `fruity`.

2. **Residuals:**
   - These are the differences between the observed and predicted values. They give you an idea of how well the model fits the data. The values range from -29.947 to 31.642.

3. **Coefficients:**
   - **Intercept (49.842):** This is the predicted value of `winpercent` when both `sugarpercent` and `fruity` are zero.
   - **sugarpercent (11.255):** For each one-unit increase in `sugarpercent`, the predicted `winpercent` increases by 11.255, assuming `fruity` is constant.
   - **fruity (-10.989):** If `fruity` is 1 (true), the predicted `winpercent` decreases by 10.989, assuming `sugarpercent` is constant.

4. **Significance Codes:**
   - '***', '**', '*': Indicate the level of significance. In this case, all coefficients are significant (very low p-values).

5. **Residual Standard Error:**
   - It's an estimate of the standard deviation of the residuals. In this case, it's 13.39.

6. **Multiple R-squared (0.1918):**
   - It represents the proportion of the variance in the dependent variable (`winpercent`) that is explained by the independent variables (`sugarpercent` and `fruity`). In this case, it's 19.18%.

7. **Adjusted R-squared (0.1721):**
   - Similar to R-squared but adjusted for the number of predictors. It penalizes the inclusion of unnecessary variables. Here, it's 17.21%.

8. **F-statistic (9.733):**
   - It tests the overall significance of the model. A higher F-statistic indicates a more significant model.

9. **p-value (0.0001612):**
   - It indicates whether the overall model is significant. In this case, the p-value is very low, suggesting that at least one predictor variable is useful in predicting `winpercent`.

Overall, the model suggests that both `sugarpercent` and `fruity` are significant predictors of `winpercent`, and the model as a whole is statistically significant. However, the R-squared values indicate that the model explains only a moderate proportion of the variability in `winpercent`.


```{r}


# Assuming 'data' is your dataframe
model <- lm(winpercent ~ bar + caramel + chocolate + crispedricewafer + fruity + hard + nougat + pluribus + peanutyalmondy + pricepercent + sugarpercent, data = candy_raw)

# Print the summary of the regression model
summary(model)

var <- c("intercept", "bar", "caramel", "chocolate", "crispedricewafer", "fruity", "hard", "nougat", "pluribus", "peanutyalmondy", "pricepercent", "sugarpercent")



```

